---
title: "Getting started with dbschemaatlas"
description: Extract, persist, inspect, and document database schemas.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with dbschemaatlas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`dbschemaatlas` keeps connection credentials, database drivers, and branding
in the consuming application. Its documentation builds entirely from the
packaged offline fixture.

## Supply a connection provider and extract metadata

A provider creates one connection whose lifecycle is owned by the extraction
call. PostgreSQL is shown here, but neither the driver nor credentials are a
package dependency.

```{r provider, eval=FALSE}
provider <- function() {
  DBI::dbConnect(
    RPostgres::Postgres(),
    dbname = Sys.getenv("ATLAS_DATABASE"),
    user = Sys.getenv("ATLAS_USER"),
    password = Sys.getenv("ATLAS_PASSWORD")
  )
}
metadata <- extract_schema_metadata(c("public", "audit"), provider)
```

Internally, connection ownership, query execution, and SQL reading are
delegated to `dbiutils`.

## Save and load metadata offline

```{r offline}
library(dbschemaatlas)
fixture <- system.file("extdata", "example-metadata", package = "dbschemaatlas")
metadata <- read_metadata(fixture)
snapshot <- file.path(tempdir(), "atlas-snapshot")
write_metadata(metadata, snapshot)
offline_metadata <- read_metadata(snapshot)
model <- build_schema_model(offline_metadata)
model_dependencies(model)
```

## Generate a schema report

```{r report, eval=FALSE}
config <- atlas_report_config(
  title = "Customer database",
  abstract = "Tables and relationships used by the customer service.",
  author = "Data engineering",
  output_directory = "schema-report"
)
render_schema_report(model, config)
generate_report_indexes(config$output_directory)
```

## Customize branding and graph colors

```{r branding, eval=FALSE}
config <- atlas_report_config(
  title = "Customer database",
  logo = "brand/logo.png",
  analytics_html = "brand/analytics.html",
  schemas = c("public", "audit"),
  schema_groups = list(application = "public", operations = "audit"),
  schema_colors = c(public = "#2563EB", audit = "#D97706"),
  filename_prefix = "customer_database",
  output_directory = "schema-report",
  domain_labels = c(public = "Application", audit = "Operations"),
  sections = c(columns = TRUE, dependencies = TRUE, usages = TRUE, graphs = TRUE),
  template_overrides = list(report = "brand/report.Rmd.tpl", style = "brand/report.css")
)
```

All overrides are explicit; neutral installed templates remain usable when no
custom resources are supplied.
